
```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE)
options(mrgsolve.soloc = "build")
```


```{r, include = FALSE}
source(here::here("src/global.R"))
library(mrgsolve)
library(tidyverse)
library(knitr)
library(lattice)
```


# Big picture

There are 3 (or 4) main workflows that we want to work up to. 

Decide what is the output you want and that will tell you what 
the inputs you need. 


## You need a model

For every workflow, you need a model. This is coded in a separate 
file and read in by `mread()`

```{r}
mod <- mread_cache("azithro-fixed.mod", project = here("model"))
```

We'll code up a model right after this overview.

## Single profile

To get a single profile, use and event object

```{r}
mod %>% 
  ev(amt = 250, ii = 24, addl = 4) %>% 
  mrgsim(end = 144, delta = 0.1) %>% 
  plot("CP") 
```


## Population

To simulate a population, you'll need what we call a data set

```{r}
mod <- mread_cache("azithro.mod", project = here("model"))
data <- expand.evd(amt = 250, WT = runif(30, 50, 100))
```

```{r, echo = FALSE}
head(data, n = 3)
```

```{r}
set.seed(9876)

mod %>% 
  data_set(data) %>% 
  mrgsim(end = 24) %>% 
  plot("CP")
```

## Batch 

You can also simulate a population (or a batch of subjects) with a 
data set of parameters and an event object


```{r}
mod <- mread_cache("azithro-fixed.mod", project = here("model"))

idata <- expand.idata(WT = runif(30, 50, 100))
```

```{r, echo = FALSE}
head(idata, n = 3)
```

```{r}

mod %>% 
  ev(amt = 250, ii = 24, addl = 4) %>% 
  idata_set(idata) %>% 
  mrgsim(end = 144) %>% 
  plot("CP")
```

## Replicate

The other pattern is just like data set, but we do that in a loop
to generate replicate simulations. 

This simulation might look something like this


```{r}
#| eval: false

sim <- function(i, model, data) {
  mod %>% 
    data_set(data) %>%
    mrgsim() %>% 
    mutate(irep = i)
}

out <- lapply(1:1000, sim, model = mod, data = data) %>% bind_rows()
```

See @sec-replicate for full examples on replicate simulation.


## The overall pattern 

- Code a model 
- Load it with `mread()`
- Set up your intervention and population
- Simulate with `mrgsim()`
- Plot or process your output

