# Data sets

```{r}
#| include: false
source(here::here("src/global.R"))
options(pillar.width = Inf)
```


`mrgsolve` will receive data sets just like those you create
for NONMEM, but in `data.frame` format. 

Naming is similar to what we learned in the previous chapter for 
event objects, but you can use upper or lower case naming. 

Example

```{r}
#| message: false
data <- read.csv(here("data/derived/analysis3.csv"), na = '.')
head(data, n = 3)
```

Data sets are passed through the `data_set` pipe function

```{r}
#| eval: false
mod %>% 
  data_set(data) %>%
  mrgsim()
```

or just pass it in as the `data` argument to `mrgsim()`

```{r}
#| eval: false
mrgsim(mod, data = data)
```


## Specifics

- Lower or upper case names may be used for NMTRAN data items, but 
  they must be consistent 
  - Good: `AMT`, `CMT`, `TIME`
  - Also good: `amt`, `cmt`, `time`
  - Bad: `amt`, `CMT`, `time`

- Columns with special handling
  - `ID`
  - `AMT`
  - `CMT`
  - `EVID`
  - `II`
  - `ADDL`
  - `RATE`
  - `SS`

- Required columns for most applications
  - `ID`
  - `TIME`
  - `CMT` 

- Some columns which are commonly used for NONMEM data sets mean 
  nothing in `mrgsolve` context
  - `C`
  - `MDV`
  - `DV`

- Columns that are not numeric will be dropped before simulating

- Use `uctran()` and `lctran()` to convert names

- Data sets can contain mix of observation and event records

- Data sets can have multiple IDs with any design
  - NONMEM approach for when to count "new" IDs

- Unlimited flexibility in creating data sets
  - But we have some helpful tools when creating simulation 
    skeletons on the fly




## Observation records

Data sets may or may not include observation records (EVID=0). 

- When there is **at least one** observation record anywhere in the 
  data set, `mrgsolve` will only take observations from the 
  input data. 
- When there **no** observation records in the data set, `mrgsolve` will
  insert observations according to the time grid specification in the 
  model object. 

This makes it easy to pass in a _clinical_ data set, with arbitrary "real"
observation times. 

Or, if you are conducting _de novo_ simulations, make a data set with only 
dosing or event records and let `mrgsolve` fill in the observations for you
after you customize the time grid. 

## Building input data sets

We will look at some convenience functions that help you create input data sets
that _don't_ include observation records. 


### Expand

The `ev_expand()` and `evd_expand()` functions like `base::expand.grid()`: it 
will make all combinations of your inputs, filling in missing required columns
with sensible defaults. 

To simulate 2 subjects at each of 3 doses

```{r}
evd_expand(amt = c(100, 300, 1000), ii = 24, total = 7, ID = 1:2)
```

- Arguments are lowercase
- The function fills in 
  - `TIME`, `CMT` and `EVID` for you; be sure to override this if you want 
    something else
- Only one row for each subject
- The result is a data.frame

### From events

Pass several event objects to `as_data_set()`. This allows much more 
flexibility. Let's do an example:

We want to simulate 2 subjects at each of 3 regimens: 

- a: 100 mg daily for 28 days
- b: 50 mg daily for 14 days followed by 100 mg daily for 14 days
- c: 50 mg BID for 14 days followed by 100 mg daily for 14 days

First, make an event object for each

```{r}
a <- ev(amt = 100, ii = 24, total = 28)
```

```{r}
b1 <- ev(amt = 50, ii = 24, total = 14)
b2 <- mutate(a, total = 28)
b <- ev_seq(b1, b2)
```

```{r}
c1 <- ev(amt = 50, ii = 12, total = 28)
c2 <- b2
c <- ev_seq(c1, c2)
```

Check the outputs

```{r}
a
b
c
```

The `as_data_set()` function will combine these into 3 different subjects, 
giving them unique ID names names

```{r}
data <- as_data_set(a,b,c)
data
```

We can expand the number of subjects in each intervention by using 
`ev_rep()`

```{r}
data <- as_data_set(
  ev_rep(a, ID = 1:2), 
  ev_rep(b, ID = 1:2), 
  ev_rep(c, ID = 1:2)
) %>% uctran()
```

```{r}
#| include: false
rownames(data) <- NULL
```


```{r}
data
```

## Individual data sets

Another type of data set is `idata`. This is a data frame of individual-level
data for the problem. 

- `idata` typically has columns named for parameters
- `idata` has only one row per subject in the problem
- `idata` defines the number of subjects in the problem when there is no
  (population) `data` set
  

For example, load the `1005` model

```{r}
mod <- modlib("1005", end = 72, delta = 0.1, capture = "WT")

see(mod)
```

This model has `WT` as a parameter (input covariate) as well as `TVKA`, `TVCL`, 
and  `TVV`.

Make an `idata` set for sensitivity analysis on `WT`

```{r}
idata <- expand.idata(WT = runif(100, 50, 110))

head(idata)
```

We have two columns

- `ID`
- `WT`

and `WT` is labeled as a input in the model

```{r}
check_data_names(idata, mod, mode = "inform")
```

Notice that `SEX` is _also_ tagged as an input and the check didn't find this 
covariate in the data set. That's ok; we're just looking at `WT` for now.


If we simulate one 100 mg dose with this `idata` set and no IIV 

```{r}
out <- 
  mod %>% 
  zero_re() %>% 
  idata_set(idata) %>% 
  ev(amt = 100) %>% 
  Req(WT, CL, IPRED) %>% 
  mrgsim()
```

we get 100 subjects in the output, all with varying weights

```{r}
out
plot(out)
```

