# Event objects

```{r}
#| include: false
library(mrgsolve)
library(dplyr)
options(mrgsolve_mread.quiet = TRUE)
```

Event objects are quick ways to generate an intervention or
a sequence of interventions to apply to your model. 

These are like quick and easy data sets. 


## Create and event object


Use `ev()` and pass NMTRAN data names in lower case. 

For example

```{r}
ev(amt = 100, ii = 12, addl = 2)
```

You can pass

- `time` time of the event
- `evid` event ID
  - 1 for dose
  - 2 for other type
  - 3 for reset
  - 4 for dose and reset
  - 8 for replace
- `amt` dose amount
- `cmt` compartment for the intervention
  - usually the compartment number
  - can be character compartment name
- `ii` inter-dose interval
- `addl` additional doses (or events)
  - `total` alternative for total number of doses
- `ss` advance to steady-state?
  - 0 don't advance to steady-state
  - 1 advance to steady-state
  - 2 irregular steady-state
- `rate` give the dose zero-order with this rate
  - `tinf` alternative for infusion time 


See `?ev` for additional details. 

## Invoke event object


### Inline

Pipe the model object to `ev()` then simulate.

```{r}
#| fig-height: 3
mod <- house(outvars = "GUT,CP,RESP", end = 24)

mod %>% ev(amt = 100) %>% mrgsim() %>% plot()
```

### As object

You can save the event object out and pass it in. 

```{r, eval = FALSE}
e <- ev(amt = 100) 

mod %>% ev(e) %>% mrgsim() %>% plot()
```

This is a good idea when you want to create an intervention 
and apply it in several different simulation scenarios.


Alternatively, you can pass it in as the `events` argument
for `mrgsim()`

```{r, eval = FALSE}
mod %>% mrgsim(events = e) %>% plot()
```

## Combining event objects 

We can create more complex interventions from several
simpler event objects. 

### Simple combination

Use the `c()` operator to concatenate. 

For 100 mg loading dose followed by 50 mg daily x6

```{r}
load <- ev(amt = 100) 
maintenance <- ev(time = 24, amt = 50, ii = 24, addl = 5)

c(load, maintenance)
```

### Sequence

We can make this simpler by putting these in a sequence. Here is 
100 mg daily for a week, followed by 50 mg daily for the rest of the month

```{r}
a <- ev(amt = 100, ii = 24, total =  7) 
b <- ev(amt = 50,  ii = 24, total = 21)

seq(a,b)
```


### Expand into multiple subjects

Pass an event object to `ev_rep()` with the IDs you want

```{r}
seq(a,b)

seq(a,b) %>% ev_rep(1:3)
```


### Combine into a data set

Use `as_data_set` with `ev_rep()` to create a data set

```{r}
c <- seq(a,b) 

as_data_set(
  a %>% ev_rep(1:2), 
  b %>% ev_rep(1:2), 
  c %>% ev_rep(1:2)
)
```

## Modifying event objects

You can use a selection of the tidyverse to modify event objects

```{r}
single <- ev(amt = 100) 

ss <- mutate(single, ii = 24, ss = 1)  

ss
```

- `mutate()`
- `select()`
- `filter()`

## Rx specification


```{r}
ev_rx("100 mg x1 then 50 q12h x 10 at 24")
```

