# The model object

```{r}
#| include: false
library(mrgsolve)
library(dplyr)
options(mrgsolve_mread.quiet = TRUE)
```


This chapter introduces the ___model object___. 

- The model object contains all information about _the model_
  - Compartments
  - ODE
  - Algebraic relationships
  - Random effects
  - Etc
- The model object is what you use in R to 
  - Query the model
  - Run simulations
  
## `mread()`

Load a model from a model specification file using `mread()`. 

- Don't worry for now what is in that file; we'll show you how to create it
- Your model can have any extension

### Syntax to load a model 

__Provide the complete path to the file__

```{r}
#| eval: false
mod <- mread("book-model/test.mod")
```


__Provide the directory as an argument__

- Assumes you are keeping all simulation code in the `models` directory

```{r}
#| eval: false
mod <- mread("test.mod", project = "book-model")
```


__Set `project` as an option__

- All my models are in this directory _for this project_

```{r}
#| eval: false
options(mrgsolve.project = "book-model")

mod <- mread("test.mod")
```

__Update on load__

- Set the simulation end time to `240`
- Set (increase) ODE solver relative tolerance to `1e-5`

```{r}
#| eval: false
mod <- mread("book-model/test.mod", end = 240, rtol = 1e-5)
```


### Read and cache

Use `mread_cache()` to build and cache the model. 

Load it once

```{r}
mod <- mread_cache("book-model/test.mod")
```

Load it again

```{r}
mod <- mread_cache("book-model/test.mod")
```

Look at the `soloc` argument; this will let you place the cached model in 
a stable location. 

## `modlib()`

Use the `modlib()` function to load a model from an internal model library.

- Pre-coded models
- Sourced from inside the `mrgsolve` installation directory
- Great way to get models to experiment with
  - But I rarely use these for production work

This code will load a 1-compartment PK model. 

```{r}
#| eval: false
mod <- modlib("pk1")
```
 
List out the location of he model library

```{r}
modlib()
```

So the `modlib()` function is equivalent to 

```{r}
#| eval: false
mod <- mread("pk1", project = modlib())
```

Find out what models are in the library

```{r}
#| eval: false
?modlib
```

## Overview

You can print `mod` to the R console and see what's going on 

```{r}
mod
```

or summarize

```{r}
summary(mod)
```

or see the model code

```{r}
see(mod)
```


## Parameters

Query the parameter list with `param()`

```{r}
param(mod)
```

This output shows you there are 3 parameters in the model 

- `CL`, with nominal value 1
- `V`, with nominal value 20
- `KA`, with nominal value 1

Note that each parameter has

- A **name**(e.g. `CL`)
- A **value** (must be _numeric_) 



## Compartments

We'll keep working with the model object we loaded in the previous section. 

Models also have compartments. Like parameters, compartments have 

- A **name**
- A **value** 

Compartments also have a ___number___ 

Query the compartment list with `init()`

```{r}
init(mod)
```

Notice that each compartment has a number associated with it. This is mainly
used for dosing. But there is a model syntax that allows you to write a 
model in terms of named compartments (e.g. `A(2)` or `F1`). 

## Random effects

```{r}
revar(mod)
```


## Update the model object

We frequently want to _change_ or _update_ the settings in the model object. 

Updates can be made through `update()`. To change the simulation time grid
we might use

```{r}
mod <- update(mod, end = 240, delta = 2)
```

to change the simulation end time to `240` hours and the output time interval
to every 2 hours. This results in a new model object with updated settings that 
will be in place whenever you simulate from `mod` until you make more changes. 

You can also update on model read

```{r}
#| eval: false
mod <- mread("model.mod", end = 240, delta = 2)
```

or at the time of simulation 

```{r}
#| eval: false
out <- mod %>% mrgsim(end = 240, delta = 2)
```

All of these update mechanisms execute updates to the model object. But only 
when we save the results back to `mod` are the updates persistent in the model.

__What else can I update?__

- Time
  - `start`, `end`, `delta`, `add`
- Parameters and compartment initial values
- Solver settings
  -  `atol`, `rtol`
  -  `hmax`, `maxsteps`, `mxhnil`, `ixpr`
  - Usually changing `rtol`, `atol`, and maybe `hmax`
- `$OMEGA`, `$SIGMA`
- `tscale` (rescale the output time)
- `digits`

__Parameter update__

To update parameters, use `param()`. More on this in @sec-parameters 

```{r}
#| eval: false
mod <- param(mod, CL = 2)
```





## Advanced

__Get the value of a parameter or setting__

```{r}
mod$CL
mod$end
```




__Extract all parameters as a list__

```{r}
as.list(param(mod))
```

Extract the value of one parameter

```{r}
mod$CL
```


__Extract everything__

You can get the model object contents as a plain list



```{r}
l <- as.list(mod)
```


::: {.callout-note collapse="true" appearance="minimal"}
# See the output

```{r}
l
```

:::