# Model parameters {#sec-parameters}

```{r}
#| include: false
source(here::here("src/global.R"))
data <- expand.grid(
  WT = c(60, 70, 80), 
  SEX = 0, 
  EGFR = 60
)
data <- mutate(
  data, 
  ID = row_number(), 
  TIME = 0,
  AMT = 100, 
  CMT = 1, 
  EVID = 1
) %>% select(ID, TIME, AMT, CMT, everything())
write.csv(
  x = data, 
  file = here("data/parameters-data.csv"), 
  quote = FALSE, row.names = FALSE
)
```


Model parameters are `name` / `value` pairs that are used _inside_ your model, 
but they can be varied _outside_ the model. 

Understanding how mrgsolve handles model "parameters" particularly important 
for generating interesting and robust simulations. 

Big picture

- `mrgsolve` maintains a parameter list, including names and values
  - This list is used by default if nothing else is done
  - The parameter values in this list can be updated

- `mrgsolve` will check input data sets for _columns_ which have the same 
  name as a parameter
  - When a match is made between data set and parameter list, `mrgsolve`
    will update the value based on what is passed on the data
  - Parameters in `idata` are checked (and parameter list  updated) first; 
    after that, the data set is checked (and parameter list updated)

## Coding model parameters

Traditionally, we've used the `$PARAM` block to set parameter names and values

```{r, eval = FALSE}
$PARAM 
WT = 70, SEX = 0, EGFR = 100
```

New in mrgsolve 1.2.0, you can use the `$INPUT` block. This is another way
to specify parameters, but they will have a special _tag_ on them that we can 
use later. 

```{r, eval = FALSE}
$INPUT 
WT = 70, SEX = 0, EGFR = 100
```


It's best if you can set these to sensible values; specifically, set to 
the _reference_ value in your covariate model.

## Updating parameter values

You can't change the name or number of parameters after you compile the model, 
but you can change the value. 

You can update parameters either 

- _prior to_ simulation or 
- _during_ simulation

We will illustrate with this model


```{r}
mod <- mread("parameters.mod", project = here("model"))
```

```{r}
param(mod)
```

There parameters are: 

- `WT`
- `SEX`
- `EGFR`
- `THETA1` ... `THETA5`

::: {.callout-note collapse=true appearance="simple"}
# See model code
```{r, code = mod@code, eval = FALSE}
```
:::

### Update _prior to_  simulation


Use `param()` to update the model object. You can do this in one of two ways. 

#### Update with `name=value`

The first way is to pass the new value with the parameter name you want to 
change. To change `WT`


```{r}
mod$WT
```

```{r}
mod <- param(mod, WT = 80)

mod$WT
```

And when we simulate, 

```{r}
mrgsim(mod) %>% plot("WT")
```

You can also do this via `update()`

```{r}
mod <- update(mod, param = list(WT = 60))

mod$WT
```

Remember that `mrgsim()` passes to `update()` so you can do the same thing
with 

```{r}
#| eval: false
out <- mrgsim(mod, param = list(WT = 70))
```

This will generate simulated output with `WT` set to 70

### Update with object

If you have a named object, you can pass that in to the update 
as well. For example, pass in a named list

```{r}
p <- list(WT = 70.2, FOO = 1)

mod <- param(mod, p)

mod$WT
```

Or a data frame

```{r}
data <- data.frame(WT = c(70, 80.1), BAR = 2)

mod <- param(mod, data[2,])

mod$WT
```

**This will be a very important pattern we'll do later on**


## Update _during_  simulation 

In this approach, we'll add a columns to our input data set
with the same names as our parameters and let `mrgsolve`
pick up the new values. 



To illustrate, load a data set from which to simulate

```{r}
data <- read.csv(here("data/parameters-data.csv"))
data
```

- Subjects 1, 2, and 3 have different (increasing) weight
- All subjects have `SEX=0`
- All subjects have `EGFR=60`


When we pass this data frame for simulation and plot

```{r}
out <- 
  mod %>% 
  data_set(data) %>%
  zero_re() %>% 
  mrgsim(delta = 0.1, end = 6)

plot(out, "WT,CP")
```



All of this _only_ works if the names in the data set
match up with the names in the model. 

## Check if the names match

Recall that we coded the model covariates using `$INPUT`, rather 
than `$PARAM`?

We can see that these parameters have this special tag

```{r}
param_tags(mod)
```

They have the `input` tag, which means we expect to find them on the data set
_when we ask_.


We can check this data set against the parameters in the model

```{r}
check_data_names(data, mod)
```


Now, modify the data set so it has `eGFR` rather than `EGFR`

```{r}
data2 <- rename(data, eGFR = EGFR)

check_data_names(data2, mod)
```



