# Simulation and outputs

```{r}
#| include: false
library(mrgsolve)
library(dplyr)
options(mrgsolve_mread.quiet = TRUE)
```

This chapter discusses

- Simulation from a model object
- Dealing with simulated output


## `mrgsim()`

Use the `mrgsim()` function to actually run the simulation

```{r}
#| message: false
mod <- modlib("pk1") %>% ev(amt = 100)
```

```{r}
mod %>% mrgsim() %>% plot()
```

### Update

The `mrgsim()` signature contains `...` which are passed to `update()`

```{r}
mod %>% mrgsim(outvars = "CP", end = 72, delta = 0.1) %>% plot()
```

### Options

There are some options that can _only_ be set when you call `mrgsim(). These
are function arguments. You can see them at `?mrgsim`

- `carry_out`: numeric data columns to _copy_ into the simulated output
- `recover`: like `carry_out` but works with any type
- `output`: pass `"df"` to get output as a regular data frame
- `obsonly`: don't return dosing records in the simulated output
- `etasrc`: should ETAs be simulated? or scraped from the data set
- `recsort`: how doses and observations having the same time are ordered
- `tad`: insert time after dose into the output
- `ss_n` and `ss_fixed`: settings for finding steady state
- `nocb`: next observation carry backward; set to `FALSE` for locf


About steady state: <https://mrgsolve.org/user-guide/steady-state.html>


### Variants

__Inputs__

There are `mrgsim()` variants which are specific to the types of inputs

- `mrgsim_e()` - just an event object
- `mrgsim_d()` - just a data set
- `mrgsim_ei()` - event + idata set
- `mrgsim_di()` - data set + idata set
- `mrgsim_i()` - just idata set


__Outputs__

You can also call `mrgsim_df()`, which is a wrapper for `mrgsim()` that 
always returns a data frame. 


__Quick__

Call `mrgsim_q()` for a quick turnaround simulation, with minimal overhead 
(and features). Only really useful when you are simulating repeatedly ... 
like when estimating parameters or doing optimal design.


## Simulated output

`mrgsim()` returns an object with class `mrgsims`; it's a data frame with some 
extra features.

```{r}
out <- mrgsim(mod)

class(out)

head(out)

summary(out)
```

### Output scope

- First column is always `ID`
- Second column is always `time`

By default, you get simulated values in all compartments and for every 
derived output _at every_ time


```{r}
head(out)
```

- `EV` and `CENT` are compartments
- `CP` is a derived variable (`CENT/V`)

```{r}
outvars(mod)
```

You can control which compartments and derived outputs are returned. 

Request specific outputs at simulation time

```{r}
mod %>% 
  update(outvars = "CP") %>% 
  mrgsim()
```

Or make the change persistent

```{r}
mod2 <- update(mod, outvars = "CP")

outvars(mod2)
```

### Copy inputs into output

Input data items can be _copied_ into the simulated output without passing 
through the model code itself. 

For most applications, use the `recover` argument to `mrgsim()`

```{r}
data <- expand.ev(amt = c(100,300)) %>% 
  mutate(dose = amt, arm = "100 mg x1", "300 mg x1")

out <- mrgsim(mod, data, recover = "dose, arm", output = "df")

count(out, dose, arm)
```

This will let you copy inputs of _any type_ into the output. 

If you just want to get numeric inputs into the output, use `carry_out`

```{r}
data <- expand.ev(amt = c(100,300)) %>%  mutate(dose = amt)

out <- mrgsim(mod, data, carry_out = "dose", output = "df")

count(out, dose)
```



## Working with `mrgsims` object

The `mrgsims` object can be convenient to work with when the output is small. 

```{r}
mod <- modlib("pk1", delta = 0.1)

out <- mrgsim(mod, ev(amt = 100))
```


```{r}
out
```


### Plot

Plot with a formula

```{r}
plot(out, CP ~ time)
```

or a vector of output names

```{r}
plot(out, "CENT CP")
```

### Filter

Use `filter_sims()`

```{r}
out2 <- filter_sims(out, time <= 5)

plot(out2)
```

## Coerce output

When output is big, these methods are less likely to be useful. 

In this case, coerce outputs to data.frame or tibble


```{r}
df <- as.data.frame(out)
df <- as_tibble(out)
head(df)
```

Once the output is coerced to data frame, it is like any other R data frame. 

Remember that you can get a data frame directly back from `mrgsim()` with 
the `output` argument

```{r}
mrgsim(mod, ev(amt = 100), output = "df") %>% class()
```

This is what you'll want to do most of the time when doing larger simulations.


### dplyr verbs

You can pipe simulated output directly to several dplyr verbs, for example
`filter()` or `mutate()`. 

```{r}
mod %>% mrgsim(ev(amt = 100)) %>% mutate(rep = 1)
```

This will first coerce the output object to a data frame and then continue
to work on the simulated data according to the functions in the pipeline.

Other verbs you can use include

- `group_by()`
- `mutate()`
- `filter()`
- `summarise()`
- `select()`
- `slice()`
- `pull()`
- `distinct()`

## Writing very large outputs to disk

See the `paquet` package: 

<https://github.com/metrumresearchgroup/paquet>
