
# Model specification 2 {#sec-specification-2}

```{r}
#| include: false
source(here::here("src/global.R"))
```


In this chapter, we'll keep going with model specification. This is the second
chapter on model specification. See @sec-specification-1 for the introductory
material.

## Covariate model

Like NONMEM, we can use `$PK` (or `$MAIN`) to code the covariate model, 
random effects, F, D, R, and ALAG, and initialize compartments. 

```{r}
#| eval: false
$PK

double CL = TVCL * pow(WT/70, 0.75) * exp(ETA(1));
```

- Any valid `c++` code is allowed
- Each line (statement) should end in semi-colon `;`


## C++ examples

```{c, eval=FALSE}
if(a == 2) b = 2;
if(b <= 2) {
  c=3;
} else {
  c=4;
}
d = a==2 ? 50 : 100;
double d = pow(base,exponent);
double e = exp(3);
double f = fabs(-4);
double g = sqrt(5);
double h = log(6);
double i = log10(7);
double j = floor(4.2);
double k = ceil(4.2);
```

Be careful of integer division

```{r}
#| eval: false
double result = 1/2; # 0
```

Here, `result` will evaluate to `0`.

Good practice to put `.0` behind whole numbers. 

```{r}
#| eval: false
double result = 1.0/2.0; # 0.5
```

Lots of help on the web
http://en.cppreference.com/w/cpp/numeric/math/tgamma


## Random effects

There are times when you _will_ need to code this manually. When estimating
with NONMEM and simulating with `mrgsolve`, these matrices will 
frequently be imported automatically via `$NMXML`.

### Omega / ETA

**Diagonal matrix**

```{r}
#| eval: false
$OMEGA
0.1 0.2 0.3
```

This is a 3x3 matrix with 0.1, 0.2, and 0.3 on the diagonal.

**Block matrix**

```{r}
#| eval: false
$OMEGA @block
0.1 0.002 0.3
```

This is a 2x2 matrix matrix with 0.1 and 0.3 on the diagonal. Sometimes it's 
easier to see when we code it like this

```{r}
#| eval: false
$OMEGA @block
0.1 
0.002 0.3
```

Random effects simulated from OMEGA are referred to with `ETA(n)`.

### Sigma / EPS

Works just like Omega / ETA, but use `$SIGMA` and `EPS(n)`.

For sigma-like theta, code it just as you would in NONMEM. 

```{r}
#| eval: false
$PARAM THETA12 = 0.025

$SIGMA 1

$ERROR
double W = sqrt(THETA12);
Y = (CENT/V) + W*EPS(1);
```

There is no `FIX` in `mrgsolve`; everything in OMEGA and SIGMA is always fixed. 

## Import estimates from NONMEM

- Use `$NMEXT` or `$NMXML`
  - `$NMEXT` reads from the `.ext` file
    - Can be faster
    - Doesn't retain `$OMEGA` and `$SIGMA` structure
  - `$NMXML` reads from the `.xml` file
    - Can be slower
    - Does retain `$OMEGA` and `$SIGMA` structure

This is the safest way to call

```{r}
#| eval: false
$NMXML
path = "../nonmem/106/106.xml"
root = "cppfile"
```

You might be able to use this `run`/`project` approach as well

```{r}
#| eval: false
$NMXML 
run = 1006
project = "../sim/"
root = "cppfile"
```

## Models in closed form

`mrgsolve` will solve one- and two-compartment models with first order
input in closed form. This usually results in substantial speed up. 
Use `$PKMODEL`. 

```{r}
#| eval: false
$PKMODEL cmt = "GUT,CENT", depot = TRUE
```

Certain symbols are required to be defined depending on the model. `mrgsolve`
models are always parameterized in terms of clearances and volumes except
for absorption, which is in terms of rate constant. 

- `CL / V`
- `CL / V / KA`
- `CL / V2 / Q / V3`
- `CL / V2 / Q / V3 / KA`

These can be defined as a parameter or a derived quantity in `$PK`.

Compartment names are user-choice; the only thing mrgsolve cares about is the 
number of compartments.

## Plugins

### autodec

Historically, you have had to _declare_ the type of any new variable you 
want to create. 

```{r}
#| eval: false
$PK
double KE = CL/V;
```

For most models, the numeric variables you declare are likely to be floating
point numbers ... with type `double`. 

We created a plugin that tells `mrgsolve` to look for new variables and 
declare them for you. 

```{r}
#| eval: false
$PLUGIN autodec

$PK
KE = CL/V;
```

### nm-vars


`mrgsolve` historically used

- `CENT`
- `dxdt_CENT`
- `F_CENT`
- `D_CENT`

etc. When we started `mrgsolve`, this was a really nice feature because you 
didn't have to think about compartment _numbers_. However, this made
translation of the model more difficult.

When you invoke the `nm-vars` plugin, you can write in a syntax that is much 
more like NONMEM.

For example

```{r}
#| eval: false
$PK
F2 = THETA(3);

ALAG2 = EXP(THETA(4));

$DES
DADT(1) = - KA * A(1);
```


Other convenience syntax 

- `LOG()` and `log()`
- `EXP()` and `exp()`
- `SQRT()` and `sqrt()`

Regardless of whether you have `nm-vars` invoked or not, you can still use
`THETA(n)` to refer to parameter `THETAn`.


Try the `nm-like` model in the model library for an example. 

```{r, eval = FALSE}
mod <- modlib("nm-like")

mod@code
```

### Rcpp (random numbers)

This gives you functions and data structures that you're used to using in R, 
but they work in `c++`. 

The main use for this is random number generation. Any d/q/p/r function 
in R will be available; arguments are the same, but omit `n` (you always 
get just one draw when calling from `c++`.

For a draw from U(0,1)

```{r}
#| eval: false
$PLUGIN Rcpp

$ERROR
double u = R::runif(0, 1);
```

Note: this will slightly increase compile time.

## Other blocks

- Use `$SETUP` to configure the model object on load
  - For example, set the simulation end time
- Use `$ENV` to define a set of R objects that might be evaluated in other 
  model blocks
- Use `$PRED` for other closed form models
- Use `$PREAMBLE` for code that gets run once at the start of a problem 
  `NEWIND==0`
- Use `$GLOBAL` to define variables outside of any other block

## Variables and macros

There is too much syntax to mention it all here. You will find all the 
syntax here

<https://mrgsolve.org/user-guide/specification.html#sec-variables-macros>


## Modeled event times

To get the model to stop at any time (even if not in the data set) with 
EVID 2

```{r}
#| eval: false
double mt1 = self.mtime(1.23 + ETA(1));
```

To get the model to stop at any time with user-specified EVID (e.g. 33)

```{r}
#| eval: false
self.mevent(1.23 + ETA(1), 33);
```
