# Validation

```{r}
#| include: false
source(here::here("src/global.R"))
library(here)
library(mrgsolve)
library(dplyr)
library(readr)
library(glue)
options(mrgsolve_mread.quiet = TRUE)
```

There are easy ways to validate the model you coded when you have a finished
NONMEM run. 

Outputs from the finished NONMEM run provide reference values and simulations
from `mrgsolve` provide test values which are compared against the reference.

__Recommendation__

- Have a dedicated script setup to validate your models or
- Put some validation code at the top of your VPC script


## Validate using `PRED`

First, load the `mrgsolve` model that you want to validate

```{r}
runno <- 106

mod <- mread(here(glue("model/pk/{runno}.mrgsolve")))
```

Next load in the model estimation data set.

```{r, message = FALSE}
csv <- read_csv(here("data/derived/analysis3.csv"), na = ".")

head(csv, n = 3)
```

Next, load the table file that contains `PRED`

```{r, message = FALSE}
tab <- read_table(here(glue("model/pk/{runno}/{runno}.tab")), skip = 1)
head(tab, n = 3)
```

Now join them together with `tab` on the left; the `NUM` column provides a key 
to join these two

```{r}
data <- left_join(tab, csv, by = "NUM")

head(data, n = 3)
```

Now, simulate from this data set after turning off all the random effects. We 
also bring `EVID` and `PRED` into the output to help post-processing

```{r}
out <- 
  mod %>% 
  zero_re() %>% 
  obsonly() %>% 
  mrgsim(data = data, recover = "NM = PRED", output = "df", digits = 5)

head(out, n = 3)
```

Notice, too, we limited the number of output digits to 5 (what NONMEM is 
giving).

Now, we have the NONMEM-generated `PRED` in the data as `NM` (the reference)
and we can pick between `IPRED` and `Y` as the test. 

```{r}
plot(out$NM, out$IPRED)

summary(out$NM - out$IPRED)
```

This `mrgsolve` model gives identical result.


## Validate using `IPRED`

More recent versions of `mrgsolve` also allow you to validate by `IPRED`. 

Read in an additional table that contain the post-hoc ETA values

```{r, message = FALSE}
tab <- read_table(
  here(glue("model/pk/{runno}/{runno}par.tab")), 
  na = '.', skip = 1
)
etas <- select(tab, matches("NUM|ETA"))
```

and we join this to the data set

```{r}
data <- left_join(data, etas, by = "NUM")
```

Now, we simulate with the `etasrc` option set to `data.all` and recovering 
`IPRED` into the output

```{r}
out <- mrgsim(
  mod, 
  data, 
  etasrc = "data.all", 
  obsonly = TRUE, 
  recover = "NMIPRED = IPRED, NMPRED = PRED", 
  digits = 5
)
```


And summarize again

```{r}
plot(out$IPRED, out$NMIPRED)

summary(100*(out$IPRED - out$NMIPRED)/out$NMIPRED)
```

