---
title: "Probability of Technical Success - EPO IV/SC QW/TIW"
---

```{r}
#| message: false
source(here::here("src/global.R"))
library(mrgsolve)
library(tidyverse)
library(future.apply)
library(future.callr)
options(future.fork.enable = TRUE)
set.seed(112)
```


__Population Pharmacokinetic and Pharmacodynamic Model- Based Comparability 
Assessment of a Recombinant Human Epoetin Alfa and the Biosimilar HX575__

- Xiaoyu Yan, MS, Philip J. Lowe, PhD, Martin Fink, PhD, 
Alexander Berghout, PhD, Sigrid Balser, PhD, and Wojciech Krzyzanski, PhD
- J Clin Pharmacol. 2012 November ; 52(11): 1624–1644. 
- https://www.ncbi.nlm.nih.gov/pubmed/22162538


# Read in data ...

```{r}
post <- readRDS(here("data/epo_post.RDS"))

post <- select(post, contains("THETA"))

head(post)

dim(post)
```

# ... and the model
```{r}
mod <- mread_cache("epo", here("model")) 
mod <- zero_re(mod)
see(mod)
```

Drop the initial `HGB` by a little bit

```{r}
mod <- param(mod,    THETA19 = mod$THETA19*0.8)
post <- mutate(post, THETA19 = THETA19*0.8)
```

# Weekly and tiw dosing

```{r}
tiw <- ev_days(ev(amt = 7000, rate = -2), days="m,w,f", addl = 3)
tiw

qw <- filter(tiw, time==0) %>% mutate(amt = 40000)
qw
```

```{r}
data_sc <- bind_rows(tiw, qw) %>% mutate(ID = amt)
data_sc
```

# SC administration

- When subcuteneous administration, 40,000 IU weekly has similar efficacy
to 100 IU/kg TIW

- We're looking out to 4 weeks here

```{r}
mod %>% 
  mrgsim(data = data_sc, end = 672) %>%
  plot("EPOi, HGBi")
```

# IV administration

```{r}
data_sc
data_iv <- mutate(data_sc, cmt = 2, rate = 0)
data_iv
```

- But when we look at the TIW versus QW comparison with IV administration,
the once-weekly is less effective than three-times-weekly

```{r}
mod %>% 
  mrgsim(data = data_iv, end = 672) %>%
  plot("EPOi, HGBi")
```

# Probability of technical success

Let's look at the probability of technical success for the QW versus TIW
dosing in the IV case

Let's ramp this up for PTS

## Population data sets

TIW data

```{r}
iv_tiw <- ev_rep(tiw, ID = 1:250) %>% mutate(cmt = 2, rate = 0)
```

QW data

```{r}
iv_qw <- filter(iv_tiw, time==0) %>% mutate(amt = 40000, ID = ID + 10000)
```

If we bind this all together, we can simulate both regimens in one go

```{r}
data <- bind_rows(iv_tiw, iv_qw) %>% mutate(qw = as.integer(amt==40000))

head(data)
```

- What about simulating some variability in Weight?


A simulation function

Arguments

- `i` replicate number
- `mod` the EPO model object
- `data` a template data set

```{r}
mod <- mread_cache("epo", here("model"), rtol = 1e-5)
mod <- param(mod, THETA19 = mod$THETA19*0.8)

sim <- function(i, mod, data) {
  
  mod <- param(mod, slice(post, i))
  
  out <- mrgsim(
    mod, 
    data = data, 
    obsonly = TRUE,
    recover = "qw", 
    end = -1, 
    add = c(0, 672), 
    output = "df"
  )
  
  mutate(out, irep = i)
}
```


Test it out

```{r}
sim(2, mod, data) %>% head()
```


## Simulate

- Remember to set a seed
- What about uncertainty in the random effect variances?

```{r}
set.seed(10020)

plan(callr, workers = 4L)

out <- future_lapply(
  seq(500),
  sim, 
  mod, data, 
  future.seed = TRUE
) %>% bind_rows()
```


## Summarize

Our target is HGB > 10

```{r}
sum1 <- 
  out %>% 
  filter(time > 0) %>%
  group_by(irep, time, qw) %>%
  summarise(success = mean(HGBi >= 10), n = n(), .groups = "drop") %>%
  mutate(week = time/168)

sum1

ggplot(sum1, aes(x = success)) + 
  geom_histogram(col = "grey") + 
  facet_grid(week ~ qw) 
```

Our success criteria is at least 75% of patients hitting the target

```{r}
sum2 <- 
  sum1 %>%
  group_by(week,qw) %>% 
  summarise(PR = mean(success > 0.75), .groups = "drop")
```

The bottom line

```{r}
sum2
```
