---
title: "Fc-OPG PK/PD in healthy post-menopausal women"
author: "Metrum Research Group"
---

```{r}
#| echo: false
source(here::here("src/global.R"))
```

# Project
```{r}
proj <- here("model")
```

## Question
- What is the median change from baseline 14 days after 3 mg/kg SC dose?
- What is the probability that median CFB is greater than 40%?

# OPG model
```{r}
mod <- mread("opg", proj)

mod
```

Checkout the random effects structure
```{r}
revar(mod)
```


# Load the simulated posterior
```{r}
post <- readRDS(here("data/opgpost.RDS")) %>% sample_n(1000)

post %>% head()

param(mod)
```


For most of our examples, we are taking output directly
from a NONMEM run. This example is a little different:
We have changed the column names from `THETA1`, `THETA2` 
etc ... to `TVCL`, `TVVC` etc ...

```{r,eval=FALSE}
head(post)

check_data_names(post, mod, tags = "tvpars")
```

```{r}
post <- mutate(post, TVVP1 = TVP1, TVVP2 = TVP2)
```

When working with results out of NONMEM, it is common to get
the `THETA` , `OMEGA` and `SIGMA` estimates all in a single 
row in a data frame.  

In order to get the information for `OMEGA` and `SIGMA` in to 
the problem, we need to go into this data and 
form matrices.

There are two specialized functions to help with this:

- `as_bmat` when the data is in block format
- `as_dmat` when the data is in diagonal format

```{r}
omegas <- as_bmat(post,"OMEGA")
sigmas <- as_bmat(post,"SIGMA")

omegas[[10]]

length(omegas)
is.list(omegas)

slice(post, 10)

omegas[[10]]
sigmas[[10]]
```


Here is a simple dosing data set to simulate 100 patients with 
3 mg/kg Fc-OPG SC x1
```{r}
sc3 <- expand.ev(ID = 1:100, amt = 3*70)

head(sc3)
```

When we do replicate simulation, it almost always pays off
to construct a function that will carry out one
replicate.


Arguments to the function are

- `i` the current simulation replicate
- `data` the dosing data set
- `des` the observation design


```{r}
sim <- function(i,data,des) {
  
  mod <- 
    mod %>%
    param(slice(post,i)) %>%
    omat(omegas[[i]]) %>% 
    smat(sigmas[[i]])
  
  mod %>%
    Req(PDDV) %>%
    mrgsim(
      data = data, 
      end = -1, 
      add = c(0, 14)*24,
      obsonly = TRUE
    ) %>% 
    mutate(irep = i) %>% 
    mutate(TVIC50 = mod$TVIC50, TVQ1 = mod$TVQ1, TVVMAX = mod$TVVMAX)
}
```


Also note 

- We update `param` with `ith` row from `post`
- We update `OMEGA` and `SIGMA` from the `ith` position 
  in the appropriate list of matrices
- We will capture the current 


It is easy to test the function.  The 10th replicate will
look like this

```{r}
sim(10, sc3, des)

slice(post, 10)
```



Simulate reps
```{r}
set.seed(22223)
# library(future.apply)
# library(future.callr)
# options(future.fork.enable = TRUE)
# plan(callr, workers = 3)
# sims <- future_lapply(1:1000, sim, data = sc3, des = des, future.seed = TRUE) 

sims <- lapply(1:1000, sim, data = sc3, des = des) 

out <- bind_rows(sims)

head(out)
```

Summarise

- First, get the baseline `NTX`
- Then calculate percent CFB


```{r}
summ <- 
  out %>%
  group_by(ID,irep) %>%
  mutate(BASE = first(PDDV), dDV = 100*(PDDV-BASE)/BASE) %>%
  ungroup()

head(summ)
```



Filter down to week 2
```{r}
summ <- filter(summ, time==336 & BASE >= 0)
```


Now, get the median
```{r}
sumi <- 
  summ %>%
  group_by(irep) %>%
  summarise(med = median(dDV))

ggplot(sumi, aes( x = med)) + geom_histogram( col = "grey")

sumi
```


## From the abstract

> Subsequent clinical trial simulations demonstrated that a single 3.0-mg/kg SC dose of Fc-OPG would be expected to produce, at 14 days post-dose, a median NTX percentage change from baseline of −45% (with a 95% prediction interval ranging from −34% to −60%)."

```{r}
ans <- signif(quantile(sumi$med, c(0.5, 0.025, 0.975)), 3)

paste0(ans[1], " (", ans[3], ",", ans[2], ")")
```

Median week-2 change from baseline with 95% interval


Probability that median cfb > 40%
```{r}
mean(sumi$med < -40)
```



Plot the distribution of the week-2 change from baseline
```{r}
ggplot(sumi, aes(x=med)) + geom_histogram(col="grey")
```


# Sensitivity analysis

We are already set up to do the sensitivity analysis

```{r}
par <- dplyr::distinct(out, irep, TVIC50, TVQ1, TVVMAX)
head(par)
sens <- left_join(sumi, par)
head(sens)
```

`TVIC50`

```{r}
ggplot(sens, aes(TVIC50, med)) + 
  geom_point() + 
  geom_smooth(method="loess",lwd=2)
```

`TVVMAX`

```{r}
ggplot(sens, aes(TVVMAX, med)) + 
  geom_point() +
  geom_smooth(method="loess",lwd=2)
```


`TVQ`

```{r}
ggplot(sens, aes(TVQ1, med)) + 
  geom_point() +
  geom_smooth(method="loess",lwd=2)
```
