---
title: "Meropenem VPC with vpc package"
author: "Metrum Research Group"
---


```{r, message = FALSE}
source(here::here("src/global.R"))
library(tidyverse)
library(mrgsolve)
library(data.table)
library(furrr)
library(vpc)
library(here)

options(future.fork.enable = TRUE)
future::plan(future::sequential)


.colSet1 <- function(x,...) scale_color_brewer(palette = "Set1", ...)
.fillSet1 <- function(x,...) scale_fill_brewer(palette = "Set1", ...)

```


# Introduction

- "Population Pharmacokinetic Analysis and Dosing Regimen Optimization of 
Meropenem in Adult Patients"
- Li et al. J Clin Pharmacol 2006
- Meropenem is broad-spectrum carbapenem antibiotic
- Efficacy related to time above MIC
- IV dosing every 8 hours by infusion or bolus
- bolus over 3 to 5 minutes
- infusion over 15 to 30 minutes
- Authors are interested in seeing if a longer infusion duration 
can increase time above MIC

# Read and fix up the data set
```{r}
data <- 
  read_csv(here("data/Simulated_DatasetMeropenem.csv"), na = '.')  %>% 
  mutate(CMT=1, DUR = AMT/RATE)
```



Derive a column that describes the infusion duration for each subject
```{r}
data %<>% 
  group_by(ID) %>% 
  mutate(DUR = first(DUR[!is.na(AMT)])) %>%
  ungroup

data <- 
  data %>%
  mutate(
    DUR = round(DUR,1), 
    RATE = replace_na(RATE,0)
  )

```

Look at distinct values of `CMT`, `EVID`, `DUR` in `data`

```{r}
count(data,CMT,EVID,DUR,RATE)
```


# Load the meropenem model

```{r}
mod <- mread("meropenem", here("model"))

```

This model looks a little different because we got it 
off of DDMoRe model repository.


A function to do the (replicate) simulation

Arguments

- `i` the replicate number

Returns simulated data set 

- time as in `des`
- `DUR` the infusion duration 
- `TIME > 0` and `Y > 0`
- Labeled with replicate number

```{r}
simvpc <- function(i,mod) {
  mod %>% 
    carry_out(DUR) %>%
    obsonly() %>%
    mrgsim_d(data,carry_out = "DV,DUR") %>% 
    filter(TIME > 0) %>%
    mutate(irep = i)
}
```

Simulate

- 300 iterations
- Use future
- Bind into a single data frame

```{r}
niter <- 300
set.seed(1234)
opt <- furrr_options(seed = TRUE)
out <- future_map_dfr(seq(niter), simvpc, mod = mod, .options = opt) 
```

This chunk needed for pred-correction
```{r}
pmod <- zero_re(mod)
set.seed(1234)
prd <- future_map_dfr(seq(niter), simvpc, mod = pmod, .options = opt) 
out[["PRED"]] <- prd[["Y"]]
```

```{r}
sims <- filter(out, Y > 0, DV > 0)
obs <- filter(sims, irep ==1)
```

The function is called `vpc`

```{r}
vpc(
  obs = obs,
  sim = sims,
  sim_cols=list(dv="Y", sim="irep"), 
  log_y = TRUE, 
  stratify = "DUR",
  ci = c(0.025, 0.975), 
  pred_corr = TRUE, 
  show = list(obs_dv=TRUE)
) 

```

