---
title: "Weight- based dosing"
---

```{r setup, include=FALSE}
source(here::here("src/global.R"))

library(mrgsolve)
library(dplyr)
library(here)
library(ggplot2)
theme_set(theme_bw() + theme(legend.position = "top"))
```


# Model


- Read in the pembrolizumab PK model

1: Elassaiss-Schaap J, Rossenu S, Lindauer A, Kang SP, de Greef R, Sachs JR, de
Alwis DP. Using Model-Based "Learn and Confirm" to Reveal the
Pharmacokinetics-Pharmacodynamics Relationship of Pembrolizumab in the
KEYNOTE-001 Trial. CPT Pharmacometrics Syst Pharmacol. 2017 Jan;6(1):21-28. 
doi: 10.1002/psp4.12132. Epub 2016 Nov 8. 
PubMed PMID: 27863143; PubMed Central PMCID: PMC5270295.


```{r}
mod <- mread(here("model/pembro-pk.mod"))
```


Check the parameters of the model

```{r}
param(mod)
```

Weight (`WT`) is an _input_ parameter

```{r}
param_tags(mod)
```

Let's do a quick simulation to check out the model

```{r}
dose <- expand.ev(amt = c(1,3,10)*78, tinf = 0.5/24)

dose
```

Simulate 

```{r}
out <- 
  mod %>% 
  zero_re() %>% 
  mrgsim_df(dose, delta = 0.1, end = 28)
```

And plot

```{r}
ggplot(out, aes(time, CP, color = factor(ID))) + 
  geom_line() + 
  scale_y_log10(limits = c(1,1000)) + 
  scale_x_continuous(breaks = seq(0,28,4))
```

Let's ramp this up to simulate a population with simulated weight

```{r}
data <- expand.ev(
  dose = c(1,3,10), 
  ID = 1:1000
) %>% uctran()
```


Simulate weights

- Weight is log-normal 
  - mean log(80)
  - sd 0.1

```{r}
set.seed(112233)

data <- mutate(
  data, 
  WT = exp(rnorm(n(), log(80), sqrt(0.1)))
)

ggplot(data, aes(x = WT)) + geom_histogram(color = "white")
```

Calculate dose amount and infusion duration

```{r}
data <- mutate(
  data, 
  AMT = dose * WT, 
  RATE = AMT / (0.5 / 24)
)
```


```{r}
check_data_names(data, mod)
```


```{r}
out <- mrgsim(
  mod, 
  data = data, 
  recover = "dose,WT", 
  end = 28, delta = 0.1, start = 0.1,
  output = "df"
)
```


```{r}
ggplot(out, aes(TIME, CP, group = ID)) + 
  facet_wrap(~dose) + geom_line() + 
  scale_y_log10() 
```


```{r}
trough <- filter(out, TIME==28)

ggplot(data = trough) + 
  geom_point(aes(WT, CP, color = factor(dose))) + 
  geom_smooth(se = FALSE, aes(WT, CP, group = dose), color = "black") 
```
